### 2016-05-01 tikz/regsim.tex
added tikz pictures

\begin{tikzpicture}
  \node[node] (m1) {$v_{1}$};
  \node[node] (m2) [right = 4cm of m1] {$v_{2}$};
  \node[node] (m3) [right = 4cm of m2] {$v_{3}$};

  \draw[edge] (m1) to node[below,very near start] {0} node[above,very near end] {1} (m2);
  \draw[edge] (m2) to node[above,very near start] {0} node[below,very near end] {0} (m3);

  \node[node] (r11) [below = 6cm of m1] {$(v_{1},0)$};
  \node[node] (r12) [above = 2cm of r11] {$(v_{1},1)$};
  \node[node] (r13) [below = 2cm of r11] {$(v_{1},2)$};

  \node[node] (l11) [below = 6cm of m3] {$(v_{3},0)$};
  \node[node] (l12) [above = 2cm of l11] {$(v_{3},1)$};
  \node[node] (l13) [below = 2cm of l11] {$(v_{3},2)$};

  \node       (anker) [below = 6.5cm of m2] {};
  \node[node] (mid0) [above right = 1cm and 1cm of anker] {$(v_{2},0)$};
  \node[node] (mid1) [below right = 1cm and 1cm of anker] {$(v_{2},1)$};
  \node[node] (mid2) [below = 3cm of anker] {$(v_{2},2)$};
  \node[node] (mid3) [below left = 1cm and 1cm of anker] {$(v_{2},3)$};
  \node[node] (mid4) [above left = 1cm and 1cm of anker] {$(v_{2},4)$};
  \node[node] (mid5) [above = 3cm of anker] {$(v_{2},5)$};

  \draw[edge] (r11) to node[left,near end] {1} node[right,near start] {0} (r12);
  \draw[edge] (r11) to node[left,near end] {0} node[right,near start] {1} (r13);
  \draw[edge,bend right] (r12.south west) to node[left,very near end] {1} node[left, very near start] {0} (r13.north west);

  \draw[edge] (l11) to node[left,near end] {1} node[right,near start] {0}(l12);
  \draw[edge] (l11) to node[left,near end] {0} node[right,near start] {1}(l13);
  \draw[edge,bend left] (l12.south east) to node[left,very near end] {1} node[left, very near start] {0} (l13.north east);

  \draw[edge] (mid0) to node[right,near start] {0} node[right,near end] {1} (mid1);
  \draw[edge] (mid1) to node[right,near start] {0} node[right,near end] {1} (mid2);
  \draw[edge] (mid2) to node[left,near start] {0} node[left,near end] {1} (mid3);
  \draw[edge] (mid3) to node[left,near start] {0} node[left,near end] {1} (mid4);
  \draw[edge] (mid4) to node[left,near start] {0} node[left,near end] {1} (mid5);
  \draw[edge] (mid5) to node[right,near start] {0} node[right,near end] {1} (mid0);

  \draw[edge] (mid0) to node[very near start, above] {2} node[very near end, above] {2} (l11);
  \draw[edge] (mid2) to node[very near start, above] {2} node[very near end, above] {2} (l13);
  \draw[edge,bend left=10] (mid4) to node[very near start, below] {2} node[very near end, above] {2} (l12);

  \draw[edge] (mid1) to node[very near start, above] {2} node[very near end, above] {2} (r11);
  \draw[edge] (mid3) to node[very near start, above] {2} node[very near end, above] {2} (r13);
  \draw[edge] (mid5) to node[very near start, below] {2} node[very near end, above] {2} (r12);

\end{tikzpicture}