### 2016-05-01 tikz/gadget.tex
first paper version to present

\begin{tikzpicture}
  \node[node] (0) {$v$};
  \node[node,above right=of 0] (1) {};
  \node[node,above left=of 0] (2) {};
  \node[node,below=of 0] (3) {};

  \draw[edge] (0) to (1);
  \draw[edge] (0) to (2);
  \draw[edge] (0) to (3);

  \node[node,below=5cm of 0] (0') {$v$};
  \node[node,above right=of 0'] (1') {};
  \node[node,above left=of 0'] (2') {};
  \node[node,below=of 0'] (3') {};

  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (1');
  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (2');
  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (3');
\end{tikzpicture}