### 2016-05-02 tikz/moveex.tex
worked on presentation

\begin{tikzpicture}
  \node[minimum size=12mm,inner sep=0mm, circle] (8) {};
  \node[node,right=of 8] (6) {};
  \node[node,right=of 6] (7) {};
  \node[node,below=of 8] (0) {};
  \node[node,below=of 6] (1) {};
  \node[node,below=of 7] (2) {};
  \node[node,below=of 0] (3) {};
  \node[node,below=of 1] (4) {};
  \node[node,below=of 2] (5) {};
  \node[node,above right=0.5cm and 1cm of 2] (8) {};
  \node[node,below right=0.5cm and 1cm of 2] (9) {};

  \draw[edge] (0) to node[very near start, above] {0} node[very near end, above] {1} (1);
  \draw[edge] (1) to node[very near start, below] {0} node[very near end, above] {1} (2);
  \draw[edge,bend right] (2.north west) to node[very near start, right] {0} node[very near end, above] {2} (0);
  \draw[edge] (0) to node[very near start, right] {1} node[very near end, left ] {0} (3);
  \draw[edge] (3) to node[very near start, above] {1} node[very near end, below] {1} (4);
  \draw[edge] (4) to node[very near start, below] {0} node[very near end, below] {1} (5);
  \draw[edge] (5) to node[very near start, above] {0} node[very near end, below] {2} (1);
  \draw[edge] (1) to node[very near start, left ] {3} node[very near end, left ] {2} (6);
  \draw[edge] (6) to node[very near start, above] {0} node[very near end, above] {1} (7);
  \draw[edge] (8) to node[very near start, above] {0} node[very near end, right] {0} (7);
  \draw[edge] (8) to node[very near start, above] {1} node[very near end, right] {2} (5);
  \draw[edge] (9) to node[very near start, below] {1} node[very near end, below] {3} (5);
  \draw[edge,bend right=20] (9) to node[very near start, below] {0} node[very near end, below] {1} (6);

  \begin{scope}[on background layer]%
      \node [fit=(0)] (layer) {};%
  \end{scope}
  \begin{scope}[on background layer]%
      \node [fit=(4)] (layer) {};%
  \end{scope}
  \only<3-3>{\begin{scope}[on background layer]%
      \node [fit=(0), fill=green!50] (layer) {};%
  \end{scope}}
  \only<4-4>{\begin{scope}[on background layer]%
      \node [fit=(3), fill=green!50] (layer) {};%
  \end{scope}}
  \only<5-5>{\begin{scope}[on background layer]%
      \node [fit=(4), fill=green!50] (layer) {};%
  \end{scope}}
  \only<6-6>{\begin{scope}[on background layer]%
      \node [fit=(3), fill=green!50] (layer) {};%
  \end{scope}}
  \only<7-7>{\begin{scope}[on background layer]%
      \node [fit=(4), fill=green!50] (layer) {};%
  \end{scope}}

\end{tikzpicture}