### 2016-05-02 tikz/portex.tex

\begin{tikzpicture}
  \node[node] (m1) {$v_{1}$};
  \node[node,right=2cm of m1] (m2) {$v_{2}$};

  \node[above left=of m1] (e1) {};
  \node[below=of m1] (e2) {};

  \node[above right=of m2] (e3) {};
  \node[below=of m2] (e4) {};

  \draw[edge] (m1) to node[very near start, above] {0} (e1);
  \draw[edge] (m1) to node[near start, left] {2} (e2);

  \draw[edge] (m2) to node[very near start, above] {1} (e3);
  \draw[edge] (m2) to node[near start, right] {2} (e4);

  \draw[edge] (m1) to node[very near start, below] {1} node[very near end, above] {0} (m2);
\end{tikzpicture}