### 2016-05-03 tikz/tapesimpesep.tex

\begin{tikzpicture}
  \node[minimum size=12mm,inner sep=0mm, circle] (8) {};
  \node[node,right=of 8,label={100:$p_{0}$}] (6) {$6$};
  \node[node,right=of 6] (7) {$7$};
  \node[node,below=of 8,label={100:$p_{3}$}] (0) {$0$};
  \node[node,below=of 6] (1) {$1$};
  \node[node,below=of 7,label={$p_{1}$}] (2) {$2$};
  \node[node,below=of 0,label={45:$p_{2}$}] (3) {$3$};
  \node[node,below=of 1] (4) {$4$};
  \node[node,below=of 2] (5) {$5$};

  \draw[move] (0) to (1);
  \draw[move] (1) to (2);
  \draw[move,bend left] (2.south west) to (0);
  \draw[move] (0) to (3);
  \draw[move] (3) to (4);
  \draw[move] (4) to (5);
  \draw[move] (5) to (1);
  \draw[move] (1) to (6);
  \draw[move] (6) to (7);
  \draw[dashed, move, decorate,
  decoration={snake,amplitude=0.5cm, segment length=3cm}, bend right=55]
  (7.north west) to (0);

  \matrix (tape) [left=of 0, table, text width=7mm, name=tape]
    { 1 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\};
  \draw[brace] (tape-1-1.south west) -- node[yshift=-6mm] {$p_{0}$} (tape-1-3.south east); 
  \draw[brace] (tape-1-4.south west) -- node[yshift=-6mm] {$p_{1}$} (tape-1-6.south east); 
  \draw[brace] (tape-1-7.south west) -- node[yshift=-6mm] {$p_{2}$} (tape-1-9.south east); 
  \draw[brace] (tape-1-10.south west) -- node[yshift=-6mm] {$p_{3}$} (tape-1-12.south east); 
\end{tikzpicture}