### 2016-05-04 presentation/main.tex
presentation

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}

\input{tikz/settings}

\usetheme{metropolis}
\metroset{numbering=fraction}

\DeclareMathOperator{\din}{\delta_{\mathit{in}}}
\DeclareMathOperator{\dout}{\delta_{\mathit{out}}}
\DeclareMathOperator{\Tid}{T_{\text{id}}}
\DeclareMathOperator{\Tsteps}{T_{\text{steps}}}
\DeclareMathOperator{\Twalk}{T_{\text{walk}}}
\DeclareMathOperator{\Thead}{T_{\text{head}}}

\title{A Grimm idea}
\subtitle{Exploring graphs with pebbles}
\author{Christoph Welzel}
\institute{Logik und Theorie diskreter Systeme, RWTH Aachen}

\begin{document}
\maketitle
\begin{frame}
  \frametitle{Motivation}
  \begin{itemize}
    \item Traversing Brobdingnagian graphs
    \item[$\rightarrow$] Web crawlers
    \item Agent moves over vertices along edges
    \item[$\Rightarrow$] Memory efficient agents
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Agents}
  \begin{quotation}
    Und als der volle Mond aufgestiegen war, so nahm Hänsel sein
    Schwesterchen an der Hand und ging den Kieselsteinen nach, die schimmerten
    wie neu geschlagene Batzen und zeigten ihnen den Weg.
  \end{quotation}
  \vspace{-0.5cm}
  \flushright{Brüder Grimm}
  \vspace{-0.5cm}
  \flushright{\emph{Hänsel und Gretel}}
  \begin{itemize}
    \item Agent carries set of markers $\rightarrow$ pebbles
    \item Leave pebbles on vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Brobdignagian graphs}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Indistinguishable vertices
      \item Bounded degree ($\Delta$)
      \item Edges can locally enumerated $\left\{0,\dots,\Delta-1\right\}$
        (Port)
      \item Agent is aware of:
        \begin{enumerate}
          \item Degree  of vertex
          \item Pebbles placed on vertex
          \item Pebbles carried by agent
          \item Port    agent entered vertex
        \end{enumerate}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Describe movement of agent by relative turns
      \item $e_{1},\dots, e_{n}$
      \item For a vertex $v$ with degree $d_{v}$ which is entered on port $p$
        is left through port $p + e_{i}\mod p$
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{theorem}
    A exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ produces a closed walk.
  \end{theorem}
  \begin{itemize}
    \item Introduce triple $(v,l,i)$ with vertex $v$ is entered
      through port $l$ and $e_{i}$ is the next turn
    \item Note that there is exactly one $(v,l,i) \leadsto (v',l',i')$
    \item Furthermore assume $t_{1} = (v',l',i') \leadsto (v,l,i)$ and
      $t_{2} = (v'',l'',i'') \leadsto (v,l,i)$:
      \begin{itemize}
        \item $t_{1}$ and $t_{2}$ both entered $v$ through $l$: $v' = v''$
        \item $t_{1}$ and $t_{2}$ lead that $i$ is the next turn: $i' = i''$
        \item $l' + e_{i'} \mod d_{v'} = l'' + e_{i''} \mod d_{v''}$ and
          $l',l''\leq d_{v'} = d_{v''}$: $l' = l''$
      \end{itemize}
    \item Only finite triples, thus projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \alt<+>{
    \begin{theorem}[Reingold]
      There is an algorithm $A$ with $\mathcal{O}(n)$ space requirement
      producing a universal exploration sequence for every regular graph with
      $n$ vertices.
    \end{theorem}
  }{
    \begin{theorem}
      Thus, there is $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$ that moves
      along a closed walk with $n$ distinct vertices (or explores the whole
      graph).
    \end{theorem}
  }
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> Establish 3-regularity
        \begin{itemize}
          \item Transform $v$ to $3\cdot d_{v}$ distinct vertices
          \item TODO: ADD BOUNDARY
        \end{itemize}
      \item<5-> Use $A$ and take first $k = 3n\cdot c$ ($c$ number of
        configurations $M_{A}$)
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<4->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate tape of size $m$ by pebbles:
      \begin{itemize}
        \item<3-> Separate into memory blocks of size $m_{1}$
        \item<4-> Represent every block by one of $\left\{p_{0},\dots,p_{\frac{m}{m_{1}-1}}\right\}$ pebbles
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \input{tikz/tapesimpebsep}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Every pebbles needs $2^{m_{1}}$ distinct \enquote{states}
        \begin{itemize}
          \item Placement on a walk with at least $2^{m_{1}}$ distinct vertices
            yields \enquote{states}
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices
    \item Identify \emph{distinct} vertices of the walk
    \item Read from and write to simulated tape
    \item Preserve tape content through steps of simulated agent
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Simulation of $M_{\text{Walk}}$}
  \begin{itemize}
    \item $M_{\text{Walk}}$ can be implemented as
      $(\mathcal{O}(1), 0, \mathcal{O}(m_{1}))$ pebble machine
    \item Used variables:
      \begin{itemize}
        \item $\Twalk$: stores $M_{\text{Walk}}$'s tape
        \item $\Tsteps$: stores number of taken steps
        \item $\Tid$: stores number of visited distinct vertices
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item Execute step
      \item Drop pebble $p_{\text{tmp}}$
      \item Store $\Twalk$ in $T'_{\text{walk}}$
      \item Restart $M_{\text{Walk}}$
      \item Execute steps until $p_{\text{tmp}}$ is observed
      \item If $\Twalk = T'_{\text{walk}}$ the vertex is new
      \item Otherwise start over
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

\begin{frame}
  \begin{center}
    \Huge Thank you for your attention.
  \end{center}
\end{frame}

\end{document}