### 2016-05-12 presentation/main.tex
slides "done"

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{mdframed}
\usepackage{graphicx}
\usepackage{appendixnumberbeamer}

\input{tikz/settings}

\usetheme{metropolis}
\metroset{numbering=fraction}
\setbeamercolor{math text}{fg=blue}

\DeclareMathOperator{\din}{\delta_{\mathit{in}}}
\DeclareMathOperator{\dout}{\delta_{\mathit{out}}}
\DeclareMathOperator{\Tid}{T_{\text{id}}}
\DeclareMathOperator{\Tsteps}{T_{\text{steps}}}
\DeclareMathOperator{\Twalk}{T_{\text{walk}}}
\DeclareMathOperator{\Thead}{T_{\text{head}}}

\title{Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles}
\subtitle{A Grimm idea}
\author{Christoph Welzel}
\institute{Logik und Theorie diskreter Systeme, RWTH Aachen}

\begin{document}
\maketitle
\begin{frame}
  \begin{center}
    \includegraphics{hanselgretel.jpg}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Outlook}
  \begin{itemize}
    \item Exploration of huge finite graphs by agents
    \item Agents are located on a vertex and move along edges
    \item Agents can drop pebbles on vertices
    \item Agent \emph{explores} a graph by systematically visiting all of
      its vertices
  \end{itemize}
\end{frame}

\section{Preliminaries}
\begin{frame}
  \frametitle{Huge graphs}
  \begin{columns}
    \column{0.55\textwidth}
    \begin{itemize}
      \item Finite, but huge
      \item Undirected
      \item Indistinguishable vertices:
      \parbox{\baselineskip}{
        \resizebox{!}{\baselineskip}{
          \begin{tikzpicture}
            \node[node] {};
          \end{tikzpicture}
        }
      }
      \item Bounded degree ($\Delta$)
      \item Edges can locally be labeled with
        \textcolor{blue}{$0,\dots,\Delta-1$} (Port)
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble machines
        \begin{itemize}
          \item $(s,p,m)$-pebble machine
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item<2-> $\din$: moving onto vertex
          \item<8-> $\dout$: leaving vertex
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> As an agent moves onto a vertex it observes:
        \begin{itemize}
          \item<4-> Port it entered through
          \item<5-> Pebbles it carries
          \item<6-> Pebbles on the vertex
          \item<7-> Degree of vertex
        \end{itemize}
      \item<9-> Computation on a vertex determines:
        \begin{itemize}
          \item<10-> Port to leave
          \item<11-> Pebbles to drop
          \item<12-> Pebbles to carry along
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.6\textwidth}
    \begin{itemize}
      \item Exploration sequence $e_{1},\dots, e_{n}$
      \item Relative movements of agent
      \item Leaving port: $\ell_{i} + e_{i}\mod d_{v}$
      \item \emph{Universal} for a class of graphs if it explores any graph of
        that class
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{0}}{\colorbox{white!0}{0}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.4\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{mdframed}
    \begin{theorem}[Reingold]
      There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
      exploration sequence for any regular graph on $n$ vertices.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \uncover<2->{
        \node[rectangle, draw=purple, below right=0.3 and 1 of d1]
        {3-Regularity};
      }
      \uncover<3->{
        \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal Lemma};
      }
      \uncover<4->{
        \node[rectangle, draw=magenta, below right=1.7 and 1 of d1]
        {Closed Walk Lemma};
      }
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
      that moves along a closed walk and either explores the graph or visits at
      least $n$ distinct vertices, for any graph with bounded degree.
    \end{theorem}
  \end{mdframed}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{mdframed}[linecolor=magenta]
    \begin{lemma}[Closed Walk Lemma]
      An agent following an exploration sequence of the form
      $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
    \end{lemma}
  \end{mdframed}
  \uncover<2->{
    \begin{mdframed}[linecolor=blue]
      \begin{lemma}[Periodic Universal Lemma]
        There exists an $\mathcal{O}(\log n)$-space algorithm producing a
        universal exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any
        $3$-regular graph with at most $n$ vertices.
      \end{lemma}
    \end{mdframed}
  }
\end{frame}

\begin{frame}
  \frametitle{$3$-regularity}
  \begin{mdframed}[linecolor=purple]
    Establish $3$-regularity
  \end{mdframed}
  \begin{center}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/regsim}}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{mdframed}
    \begin{theorem}[Reingold]
      There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
      exploration sequence for any regular graph on $n$ vertices.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal Lemma};
      \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed Walk Lemma};
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
      that moves along a closed walk and either explores the graph or visits at
      least $n$ distinct vertices, for any graph with bounded degree.
    \end{theorem}
  \end{mdframed}
\end{frame}

\section{Pebble simulation}
\begin{frame}
  \begin{mdframed}
    \begin{theorem}
      There is a constant $c\in\mathbb{N}$, such that for any graph $G$
      with bounded degree and any $(s,p,2m)$-pebble machine $\mathcal{M}$,
      there exists a $(cs,p+c,m)$-pebble machine $\mathcal{M}'$ that simulates
      the walk of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  \end{mdframed}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate $m$ bits by pebbles:
      \begin{itemize}
        \item<3-> Separate into $m_{1}$ sized blocks
        \item<4-> Represent blocks by pebbles
          $p_{0},\dots,p_{\frac{m}{m_{1}}-1}$
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \input{tikz/tapesimpebsep}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Pebble needs $2^{m_{1}}$ distinct \enquote{states}
      \item State as index of closed walk
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Mechanisms of simulation}
  \begin{enumerate}
    \item Simulation of $M_{\text{walk}}$ providing
        walk with $2^{m_{1}}$ distinct vertices
    \item \textcolor<2->{blue}{Identify \emph{distinct} vertices of walk}
    \item Read from and write to simulated tape
    \item Preserve tape content through steps of simulated agent
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{2. Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item \textcolor<3,8>{magenta}{Execute step of $M_{\text{walk}}$}
      \item \textcolor<4,9>{magenta}{Drop pebble $p_{\text{tmp}}$}
      \item \textcolor<4,9>{magenta}{Store $\Twalk$ in $T'_{\text{walk}}$}
      \item \textcolor<5,10>{magenta}{Restart $M_{\text{walk}}$}
      \item \textcolor<6,11-18>{magenta}{Execute steps of $M_{\text{walk}}$
          until $p_{\text{tmp}}$ is observed}
      \item \textcolor<19-20>{magenta}{If $\Twalk = T'_{\text{walk}}$ the
          vertex is new (increase $\Tid$)}
      \item \textcolor<7>{magenta}{Otherwise start over}
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Simulation of a pebble machine}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\textcolor{green}{{\checkmark}}}
    \item Identify \emph{distinct} vertices of the walk
      \uncover<3->{\textcolor{green}{{\checkmark}}}
    \item Read from and write to simulated tape \uncover<4->{\textcolor{green}{{\checkmark}}}
    \item Preserve tape content through steps of simulated agent
      \uncover<5->{\textcolor{green}{{\checkmark}}}
  \end{itemize}
  \uncover<6->{
    \begin{mdframed}
      \begin{theorem}
        There is a constant $c\in\mathbb{N}$, such that for any graph $G$
        with bounded degree and any $(s,p,2m)$-pebble machine $\mathcal{M}$,
        there exists a $(cs,p+c,m)$-pebble machine $\mathcal{M}'$ that simulates
        the walk of $\mathcal{M}$ or explores $G$.
      \end{theorem}
    \end{mdframed}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{mdframed}
    \begin{theorem}
      There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
      that moves along a closed walk and either explores the graph or visits at
      least $n$ distinct vertices, for any graph with bounded degree.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=1.5 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[below right=0.6 and 1 of d1, rectangle,draw=black] {Apply simulation Theorem $\log\log n$ times};
    \end{tikzpicture}
  \end{center}
  \begin{mdframed}
    \begin{theorem}
      Any bounded-degree graph on at most $n$ vertices can be explored using
      $\mathcal{O}(\log\log n)$ pebbles and memory.
    \end{theorem}
  \end{mdframed}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct $3$-regular graphs which an agent $\mathcal{A}$ with $p$
      pebbles cannot traverse
    \item Called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item For every pair $(a,b)$ in
      $\left\{v_{1},v_{2}\right\}\times\left\{v_{3},v_{4}\right\}$
      $\mathcal{A}$ cannot traverse a $p$-barrier from $a$ to $b$
      or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Locality gadget}
  \begin{itemize}
    \item Replace edge by $(r-1)$-barrier
    \item Enforce \enquote{locality} of pebbles
      \begin{center}
        \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/locgadget}}}
      \end{center}
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{mdframed}
    \begin{theorem}
      Given an $(r-1)$-barrier $B$ with $m$ vertices for an agent $\mathcal{A}$
      with $p\geq r$ pebbles, we can construct an $r$-barrier $B'$ with
      $\mathcal{O}(\binom{p}{r}\cdot m\cdot\alpha_{B}^{2})$ vertices for
      $\mathcal{A}$.
    \end{theorem}
  \end{mdframed}
  where $\alpha_{B}$ is the amount of \enquote{local} configurations for
  $\mathcal{A}$ with $r$ pebbles
\end{frame}

\begin{frame}
  \frametitle{Barrier construction: Idea}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Subset of pebbles of cardinality $r$
      \item Locality gadget as edges
      \item Pebbles are local: configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction}
  \begin{itemize}
    \item $B$ has no pebbles and cannot traverse $0$-barrier
    \item $r$-barrier for chosen subset of pebbles
    \item Connect barriers for all possible subsets of pebbles with cardinality
      $r$:
  \end{itemize}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{\input{tikz/iterbarrier}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Barrier base case}
  \begin{mdframed}
    \begin{theorem}[Fraigniaud et al.]
      For any $q$ non cooperative $s$-state agents without pebbles, there
      exists a $3$-regular graph $G$ on $\mathcal{O}(qs)$ vertices with the
      following property: There are two edges $\left\{v_{1},v_{2}\right\}$
      and $\left\{v_{3},v_{4}\right\}$ in $G$, the first labeled 0, such
      that any of the $q$ agents starting in $v_{1}$ or $v_{2}$ does not
      traverse the edge $\left\{v_{3},v_{4}\right\}$.
    \end{theorem}
  \end{mdframed}
  \begin{itemize}
    \item For agent $\mathcal{A}$ consider
      $S = \left\{\mathcal{A}_{q}\mid q\in Q\right\}$
    \item Since $|S| = |Q| = q$ there is a 0-barrier with $\mathcal{O}(s^{2})$
      vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier}
  \begin{mdframed}
    \begin{theorem}
      Given an $(r-1)$-barrier $B$ with $m$ vertices for an agent $\mathcal{A}$
      with $p\geq r$ pebbles, we can construct an $r$-barrier $B'$ with
      $\mathcal{O}(\binom{p}{r}\cdot m\cdot\alpha_{B}^{2})$ vertices for
      $\mathcal{A}$.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \resizebox{0.7\textwidth}{!}{\input{tikz/iterbarrier}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Lower pebble bound}
  \begin{mdframed}
    \begin{theorem}
      For any constant $\epsilon > 0$, an agent with at most
      $\mathcal{O}((\log n)^{1-\epsilon})$ bits of memory needs at least
      $\Omega(\log\log n)$ distinguishable pebbles for exploring all
      graphs on at most $n$ vertices.
    \end{theorem}
  \end{mdframed}
\end{frame}

\section{Conclusion}
\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
    \item Exploring graphs with $\mathcal{O}(\log\log n)$ pebbles and memory
    \item Simulation via pebbles
    \item Limitations of pebbles: Barrier construction
    \item Every agent with sub-logarithmic memory needs $\Omega(\log\log n)$
      pebbles to explore $n$ vertices
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{center}
    \Huge Thank you for your attention.
  \end{center}
\end{frame}

\appendix
\begin{frame}
  \frametitle{1. Simulation of $M_{\text{walk}}$}
  \begin{itemize}
    \item $M_{\text{Walk}}$ implemented as
      $(\mathcal{O}(1), 0, \mathcal{O}(m_{1}))$-pebble machine
    \item Used variables/pebbles:
      \begin{itemize}
        \item $\Twalk$: stores $M_{\text{walk}}$'s tape
        \item $\Tsteps$: stores number of taken steps
        \item $\Tid$: stores number of visited distinct vertices
        \item $p_{\text{start}}$: marks beginning of walk
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{3. Memory management}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item $\Thead$: head position
      \item $p_{\left\lfloor\frac{\Thead}{m_{1}}\right\rfloor}$: pebble of
        memory block
      \item $\mathit{off} = \Thead\mod m_{1}$: offset in memory block
      \item Reading:
        \begin{itemize}
          \item Retrieve encoding pebble
          \item Return $\mathit{off}$-th bit
        \end{itemize}
      \item Writing:
        \begin{itemize}
          \item Read bit
          \item On bit-flip move encoding pebble $2^{\mathit{off}}$ bits up or
            down
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimpebsep}}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{3. Memory management: Writing}
  \begin{itemize}
    \item Read bit
    \item On bit-flip: move pebble
  \end{itemize}
  \resizebox{\textwidth}{!}{\input{tikz/bitwrite}}
\end{frame}

\begin{frame}
  \frametitle{4. Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated agent moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex $v$
        \item $\omega'$ walk of $M_{\text{walk}}$ starting from $v$
        \item Transfer encoding pebbles onto $\omega'$
        \item Place pebbles at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier size}
  \begin{mdframed}
    \begin{theorem}
      For $r\leq p$ and $s\geq 2^{p}$, the number of vertices of an $r$-barrier
      $B_{r}$ for the $s$-state agent $\mathcal{A}$ with $p$ pebbles is bounded
      by $\mathcal{O}(s^{8^{r+1}})$.
    \end{theorem}
  \end{mdframed}
  \begin{center}
    \resizebox{\textwidth}{!}{\input{tikz/barriersize}}
  \end{center}
\end{frame}

\end{document}