### 2016-05-12 tikz/locgadget.tex

\begin{tikzpicture}
  \node[node] (0) {$v$};
  \node[node,above right=of 0] (1) {};
  \node[node,above left=of 0] (2) {};
  \node[node,below=of 0] (3) {};

  \draw[edge] (0) to (1);
  \draw[edge] (0) to (2);
  \draw[edge] (0) to (3);

  \node[node,right=5cm of 0] (0') {$v$};
  \node[node,above right=of 0'] (1') {};
  \node[node,above left=of 0'] (2') {};
  \node[node,below=of 0'] (3') {};

  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (1');
  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (2');
  \draw[edge,on background layer] (0') to node[draw,midway,fill=gray,shape=rectangle] {$B$} (3');

  \draw[->, shorten <=0.2cm, shorten >=0.2cm,
  decorate,decoration={snake,amplitude=0.3cm, segment length=1.0cm}] (0) to (0');

  \only<3->{
    \begin{scope}[on background layer]
      \node[right=4.5cm of 0, circle, fill=red!50, minimum size=22mm] {};
    \end{scope}
  }
\end{tikzpicture}