### 2016-05-12 tikz/portex.tex

\begin{tikzpicture}
  \node[node] (m1) {};
  \node[node,right=2cm of m1] (m2) {};

  \node[above left=of m1] (e1) {};
  \node[below=of m1] (e2) {};

  \node[above right=of m2] (e3) {};
  \node[below=of m2] (e4) {};

  \draw[edge] (m1) to node[very near start, above] {\textcolor{blue}{0}} (e1);
  \draw[edge] (m1) to node[near start, left] {\textcolor{blue}{2}} (e2);

  \draw[edge] (m2) to node[very near start, above] {\textcolor{blue}{1}} (e3);
  \draw[edge] (m2) to node[near start, right] {\textcolor{blue}{2}} (e4);

  \draw[edge] (m1) to node[very near start, below] {\textcolor{blue}{1}} node[very near end, above] {\textcolor{blue}{0}} (m2);
\end{tikzpicture}